---
title: "ENVS-193DS-spring2026-Final"
author: "Mila Matich"
date: "2026-06-09"
format:
  pdf:
    toc: true # use this to display a table of contents
execute:
  message: false # use this to make sure messages don't show up
  warning: false # use this to make sure warnings don't show up
---

[Link to my Github repo](https://github.com/themilamatich/ENVS-193DS_spring-2026_final)

# Set Up 
```{r}
#| label: set-up-load-packages-read-in-data

# installing packages needed 
install.packages("DHARMa")

# loading in packages for final 
library(tidyverse) # general use
library(janitor) # 
library(here) # file organization
library(MuMIn) # model selection
library(DHARMa) # cleaning data frames
library(ggeffects) # getting model predictions

# reading in data and storing as objects 
nests <- read_csv(here("data","nest_data_final.csv"))
personal <- read_csv(here("data","personal_data.csv"))

```


# Problem 1: Research Writing 

## a. Transparent statistical methods

**In part 1, they used a generalized linear model (GLM)** to how distance to water edge predicts the probability of American Avocet microhabitat use. **In part 2, they used a chi-sqaured test** to compare counts between bird species across different habitat sites.

## b. Suggestions for rewriting

#### Part 1: 

Habitat distance to water edge was a significant predictor of whether or not the shorebird American Avocet used the microhabitat. 


#### Part 2: 

Our study shows that the distribution of bird species differed significantly across different habitat types (managed pond, non-tidal marsh, salt pond, tidal marsh, or tidal mudflat). These results suggests that the studies bird species (American Avocet, Forster’s Tern, Black-necked Stilt) use habitats differently, and may prefer some habitats more than others. 


## c. Further analysis

One additional variable that could influence the probability of American Avocet microhabitat use could be seasonality measured across a long time series and keeping track of date (categorical variable) that would be categorically placed into Fall, Spring, Summer or Fall. This factor would influence the habitat use as it influences shorebird migration patterns, which determines whether or not the birds are in the location of study.

# Problem 2: Data Analysis

## a. Clean and wrangle 
```{r}
#| label: cleaning-and-wrangling-data


```

## b. Response variable 

## c. Purpose of study 

## d. Table of models 

## e. Run the models 

## f. Select the best model 

## g. Check the diagnostics 

## h. Visualize the models predictions 

## i. Write a caption for your figure 

## j. Calculate model predictions 

## k. Interpret your results 

# Problem 3. Affective and exploratory visualizations

## a. Comparing visualizations

## b. Designing an analysis

## c. Check any assumptions and run your analysis

## d. Create a visualization

## e. Write a caption 

## f. Write about your results 

## g. Sharing your affective visualization 


